### 2025-02-24 paper/relatedwork.tex
More writing.

@@

\paragraph{Typed rank polymorphism}
Gibbons showed how to express a form of rank polymorphism in Haskell,
through the use of \emph{Naperian
  functors}~\cite{10.1145/2976022.2976023}. This form of rank
polymorphism is somewhat more limited than ours; in particular, it
imposes static ranks. Further, Gibbons' work focused solely on the
embedding in Haskell, while our system also supports code extraction
and demonstrates good real-world performance.

>>>>

In the following we relate the specifics of our contribution to prior
work. One important distinction however, which is easily obscured by
this detail-oriented comparison, is that our work combines many parts
that have been subject to prior work, but are not commonly found
together.

@@

\paragraph{Type systems}
Verifying index operations is a classic application of dependent
types, as seen in for example Dependent
ML~\cite{10.1145/292540.292560}, although most work assumes static
ranks. Our target language, Futhark, supports size-dependent
types~\cite{10.1145/3609024.3609412}, but this information is used
only to ensure shape conformance (e.g., that the operands to a matrix
multiplication have the same size), not to ensure the correctness of
array indexing.

Gibbons showed how to express rank polymorphism in Haskell, through
the use of \emph{Naperian functors}~\cite{10.1145/2976022.2976023}.
This form of rank polymorphism is somewhat more limited than ours; in
particular, it imposes static ranks. Further, Gibbons' work focused
solely on the embedding in Haskell, while our system also supports
code extraction and demonstrates good real-world performance.
